### 2013-09-18 chi2014/results.tex
New results section

Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees (Table~\ref{tab:idea_forest}). In general, category trees were shallow, with a maximum tree height of 5. The number of nodes within a category tree ranged from 1 to 66, with the distribution of values roughly following a Poisson distribution. The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with number of instances ranging from 1 to 63.
\begin{table}
	\begin{tabular}[h!]{r | l l l l l l l}
	\textbf{number condition} & 5 & 10 & 20 & 50 \\ \hline \hline
	HITs & 57 & 47 & 23 & 10\\
	instances & 293 & 471 & 453 & 500 \\
	ideas & 171 & 249 & 278 & 341 \\
	category trees & 72 & 79 & 93 & 114 \\
	non-singleton trees & 28 & 34 & 40 & 48 \\ \hline \hline
	\textbf{number condition} & 75 & 100 & all \\ \hline \hline
    HITs & 10 & 10 & 146\\
	instances & 634 & 855 & 3007\\
	ideas & 443 & 177 &1212\\
	category trees & 172 & 177 &321\\
	non-singleton trees & 49 & 61 &92\\ 
	\end{tabular}
    \caption{Result counts between conditions}
    \label{tab:idea_forest}
\end{table}

@@

\subsection{Rates of Idea and Category Production}
The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., the notion that rates of new ideas and categories of ideas will diminish over time, and will do so as a function condition).

Figure~\ref{fig:cumulative_ideas} shows the cumulative count of new ideas as a function of the number of instances gathered in the experiment. For reference, a perfect ``1:1'' idea generation line is also plotted.

We model this quantity using the following logarithmic model:
\[n_{ideas} = b_0 + b_{cond} * \log (x_{scale} * x + x_{offset})\]
where $n_{ideas}$ represents the number of new ideas, $b_0$, $x_{scale}$, and $x_{offset}$ are parameters shared between conditions that serve to position the curve, and $b_{cond}$ acts as a per-condition scaling factor. This model does not explicitly model variability due to different workers, a point we return to later.

We fit the parameters to the data using a Bayesian approach \cite{kruschke_doing_201}, using the STAN modeling software to perform the actual fit \cite{stan-software:2013}. In modeling the parameters, we use uninformative priors. The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not. In our analyses, we use an HDI of 95\%.

Analysis of the HDIs of $b_{cond}$ reveals no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions. [TODO: Table of fitted values and their HDIs.] Notable in our data is the 75 condition, which rises above the 100 condition in this plot. Further analysis reveals that a single worker is responsible for that additional rise in number of new ideas. Removing that worker from the data reveals that the 75 condition is not different from either the 50 or 100 conditions.

Taking the derivative of the models allows us to examine the same data, but as rates at which new ideas are produced. As expected, rates of new ideas quickly taper off as a function of the number of ideas received.

@@

We can perform a similar analysis for the number of new idea categories that arrive over time. As would be surmised from the previous graphs, we observe a similar pattern, which can also be modeled using a logarithmic model. In the fitted model, we see that the number of new idea categories produced increases more slowly than the number of new ideas. This finding suggests that although individuals may produce new, unique ideas over time, these new ideas are within the \emph{same} categories. In other words, they are ``riffing'' -- coming up with permutations on existing ideas.

Collectively, these results largely affirm hypotheses 1 and 2: The rates of new idea and new idea categories production quickly taper off and tend toward zero over time. Furthermore, we see differences in rates between conditions. In our data, we see three clear groups of rates: the 5 and 10 conditions, the 20 condition, and the 50, 75, and 100 condition.

@@

\subsection{Characteristics of an Idea Within a Run}
We now turn our attention to individual runs, and characterize the ideas produced throughout a run.

\subsubsection{Most Popular Ideas}
Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas. This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.

@@

In both instances, we observe a steady increase in o-scores for the first 20-30 responses. After these first 20-30 responses, the o-scores essentially plateau, with no clear discernible pattern for either ideas or idea categories.

These graphs lend support for hypothesis 3, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.

To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total). If our intuition is correct, these ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances. We introduce two Bernoulli random variables to model this. The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$). The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} ($\textgreater5$) instances in a run ($P(\theta_l|\overline{top})$).

We fit these parameters using data from all runs that have more than 10 responses. We chose this constraint so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$. The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342). This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run.

@@

The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order}) also supports hypothesis 4, suggest by Parnes et al \cite{parnes_effects_1961}, that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half.
We test a stronger guarantee implied by our data: that ideas after the first 20 have higher o-scores than the first 20.
Although we expected the o-scores to be normally distributed in each condition, they were not. Instead, we fit Beta distributions to o-score for each of the $\leq20$, $\textgreater20$ conditions. Table~\ref{tab:hyp4} and Figure~\ref{fig:idea_oscore_hyp4} show the resulting parameters and models.

\begin{table}
	\begin{tabular}[h!]{r | l l }
	    & $\alpha$ median & $\alpha$ 95\% HDI \\ \hline \hline
        idea o-score (\textless=20) & 135.86 & 122.39-149.72 \\
        idea o-score (\textgreater20) & 259.84 & 240.47-278.81 \\ \hline \hline
	    & $\beta$ median & $\beta$ 95\% HDI \\ \hline \hline
        category o-score (\textless=20) & 1.16 & 1.08-1.25 \\
        category o-score (\textgreater20) & 0.82 & 0.77-0.86 \\
	\end{tabular}
    \caption{Hypothesis 4 Beta models}
    \label{tab:hyp4}
\end{table}

@@

The $\alpha$ parameter is significantly different in both the idea and category o-score comparison. The mean idea o-score is 0.997 in the $\textgreater20$ versus 0.995 in the $\leq20$ condition. Similarly, the $\leq20$ condition is favourable in category o-score with a mean of 0.982 vs 0.975.

This support the hypothesis, it provides a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for less from more.

\subsection{Idea Generation Patterns}
Given the clear differences in characteristics of ideas as a function of response number, we now consider idea generation patterns observable from the data. In this section, we examine these patterns in the context of hypotheses 5 and 6. Specifically, these hypotheses draw upon the SIAM model's conceptualization of brainstorming as a process in ideas are generated within one category image at a time.

In our data, we identify this pattern of creating new instances based on old ones as ``riffing'' (Table~\ref{tab:riffing}). The higher the number of instances requested, the more likely participants were to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas). However, these riff sequences are typically quite small in length, across conditions.

@@

As an example of riffing, below is an excerpt of a particular participant's brainstorming run.

The participant first touches on the general idea of recycling. Then, they become more specific by reducing the scope of recyling to specific parts. Finally, they limit the scope of parts to precious metals, particularly gold. At this point, the participant exhausts their working memory of the category an moves on.

SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5). Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).

@@

To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree.

To explicitly test hypothesis 5, we model the probability of any idea category being followed by the same idea category with a Bernoulli random variable. Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (95\% HDI 14.3\% - 17.0\%). To establish a baseline to understand whether this probability is greater than chance, the most popular category tree in our data set is \emph{music player}, representing 5.65\% of all instances. Thus, the greatest probability of one instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI. Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.

@@

To test hypothesis 6, we now examine the time it takes to generate a response instance. We calculate this as the time from the first activation of the form widget to the last.

An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data. We fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.

@@

The resulting fit models are shown in Figure~\ref{fig:hyp5_comparison}, overlapping the log-space histograms for observed values. The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3). The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively. These difference were both sigificant. Within-category instance generation took on average 12.6 seconds less than between-categories. This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between conditions. Thus, we find support for hypothesis 6 and also show the applicability of the SIAM model to brainstorming in microtask marketplaces.

>>>>

Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees (Table~\ref{tab:idea_forest}). The number of nodes within a category tree ranged from 1 to 66, with the distribution of values roughly following a Poisson distribution. The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with the number of instances ranging from 1 to 63.
\begin{table}
	\begin{tabular}[h!]{r | l l l l l l l}
	\textbf{number requested} & 5 & 10 & 20 & 50 \\ \hline \hline
	HITs & 57 & 47 & 23 & 10\\
	instances & 293 & 471 & 453 & 500 \\
	ideas & 171 & 249 & 278 & 341 \\
	category trees & 72 & 79 & 93 & 114 \\
	non-singleton trees & 28 & 34 & 40 & 48 \\ \hline \hline
	\textbf{number requested} & 75 & 100 & all \\ \hline \hline
    HITs & 10 & 10 & 146\\
	instances & 634 & 855 & 3007\\
	ideas & 443 & 177 &1212\\
	category trees & 172 & 177 &321\\
	non-singleton trees & 49 & 61 &92\\ 
	\end{tabular}
    \caption{Result counts between conditions}
    \label{tab:idea_forest}
\end{table}

@@

\subsection{Rates of Idea and Category Production}
Using the fully clustered data set, we can model the rate at which unique ideas and unique categories of ideas are produced, testing hypothesis 1 (i.e., the notion that rates of unique ideas and categories of ideas are non-linear over time).

Figure~\ref{fig:cumulative_ideas} shows the cumulative count of unique ideas per condition, as a function of the number of instances gathered in the experiment. The plots for each condition were obtained by considering each response instance in the order in which it was received, with its uniqueness assessed relative to all response instances received before it. For reference, an ideal ``1:1'' idea generation line is also plotted.

We model this quantity using the following exponential model:
\[n_{ideas} = y_{scale} * x^{rate}\]
where $n_{ideas}$ represents the number of unique ideas, $y_{scale}$ serves to scale the model, $x$ is the response number, and $rate$ models the rate of novel ideas. Importantly, this simplified model does not model either the condition or the effects of workers. In our analysis of the data, we found that individual workers introduced considerable variation, preventing us from separating the effects of the condition from the worker. Accordingly, we only consider this simplified model.

Fitting this model with Stan, we obtain a $y_{scale}$ value of 1.49 (HDI 1.44-1.55) and a $rate$ value of 0.87 (HDI 0.86-0.88). Because $rate$'s HDI does not include 1, we find support for hypothesis 1: the number of unique ideas is non-linear as a function of the number of responses received. The fitted model is shown as the ``Best Fit'' line in Figure~\ref{fig:cumulative_ideas}.

@@

We use an identical model for modeling the rate at which new idea categories are introduced over time. Figure~\ref{fig:cumulative_categories} shows the data and fitted model for the cumulative number of novel categories over time. For this model, we obtain a $y_{scale}$ value of 2.1 (HDI 2.0-2.2), with a $rate$ of 0.67 (HDI 0.66-0.67). Again, we find very clear support that the rate of novel categories is non-linear as a function of the number of responses received.

The visibly lower rate of growth for novel categories growth suggests workers are more likely to produce new ideas within existing categories, rather than create both novel ideas that also represent novel categories. We more closely examine characteristics of individuals' brainstorming runs next.

@@

\subsection{Characteristics of an Idea Within a Run}

\subsubsection{Most Popular Ideas}
Our data suggest that individuals typically produce ideas from a common set of idea categories early in the brainstorming process, then diverge and produce more original ideas. This phenomenon can be observed by examining the o-scores of ideas and idea categories.

Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} show the o-scores for response instances and idea categories, within a sliding window of size 10. Each point represents the mean o-score within that sliding window, with standard error bars provided for each point.

@@

In both cases, we observe a steady increase in o-scores for the first 20-30 responses. After these first 20-30 responses, the o-scores plateau, with no clear discernible pattern for either ideas or idea categories.

These graphs lend support for hypothesis 2, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.

To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total). If individuals are more likely to initially generate the same set of ideas, the top 5\% of the ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances. We introduce two Bernoulli random variables to model this. The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run (i.e., $P(\theta_f|top)$). The second, defined by parameter $\theta_l$, is the probability that an instance is in the top \% given it is one of the \emph{last} ($\textgreater5$) instances in a run (i.e., $P(\theta_l|\overline{top})$).

We fit these parameters using data from all runs that have more than 10 responses. We chose this constraint so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$. Using an uninformative beta prior, the resulting parameters are $\theta_f = 0.56$ (95\% HDI 0.52-0.61) and $\theta_l = 0.32$ (95\% HDI 0.30-0.34). This difference is significant, supporting the hypothesis that common ideas are over-represented in the early part of a brainstorming run.

@@

The increase in o-score over time in Figures~\ref{fig:idea_oscore_order} and \ref{fig:cat_oscore_order} also supports hypothesis 3, the notion that ideas generated in the second half a brainstorming session have higher o-scores than those in the first half. We test a stronger guarantee of this hypothesis that is implied by our data: that ideas after the first 20 have higher o-scores than the first 20.

Partitioning the data into instances $\leq20$ and $\textgreater20$, we found that these distributions of o-scores are best modeled using beta distributions. Figures~\ref{fig:idea_oscore_hyp4} and \ref{fig:cat_oscore_hyp4} show the data and fitted models. Examining the HDIs of the resultant model parameters shows significant differences between the models, confirming hypothesis 3.

Collectively, these results provide a clear guideline for maximizing originality in brainstorming tasks on microtask marketplaces: it is more cost-effective to solicit $\textgreater20$ ideas from a smaller pool of participants than to ask for fewer ideas from more individuals.

%The $\alpha$ parameter is significantly different in both the idea and category o-score comparison. The mean idea o-score is 0.997 in the $\textgreater20$ versus 0.995 in the $\leq20$ condition. Similarly, the $\leq20$ condition is favourable in category o-score with a mean of 0.982 vs 0.975.

% \begin{table}
% 	\begin{tabular}[h!]{r | l l }
% 	    & $\alpha$ median & $\alpha$ 95\% HDI \\ \hline \hline
%         idea o-score (\textless=20) & 135.86 & 122.39-149.72 \\
%         idea o-score (\textgreater20) & 259.84 & 240.47-278.81 \\ \hline \hline
% 	    & $\beta$ median & $\beta$ 95\% HDI \\ \hline \hline
%         category o-score (\textless=20) & 1.16 & 1.08-1.25 \\
%         category o-score (\textgreater20) & 0.82 & 0.77-0.86 \\
% 	\end{tabular}
%     \caption{Hypothesis 4 Beta models}
%     \label{tab:hyp4}
% \end{table}

@@

\subsection{Idea Generation Patterns}
Given the clear differences in idea characteristics as a function of response number, we now consider idea generation patterns observable from the data. In this section, we examine these patterns in the context of hypotheses 4 and 5, the hypotheses that draw upon the SIAM model of brainstorming. This model predicts individuals will generate ideas within ``images'' or categories of ideas (i.e., ``riff'' on an idea or concept) before switching to a new category.

In our clustered data, we can identify riffing by examining whether a response instance belongs to the same category tree as the previous instance. We can then derive a number of statistics from these data.

To contextualize this process, below is an excerpt of a particular participant's brainstorming run.

The participant first touches on the general idea of recycling, then becomes more specific by reducing the scope of recyling to specific parts. Finally, they limit the scope of parts to precious metals, particularly gold. At this point, the participant exhausts their working memory of the category and moves on.

Table~\ref{tab:riffing} provides a summary of riffing statistics observed in our data set. We find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas. However, we observe that riff sequences are typically quite small in length, across conditions.

@@

SIAM predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 4). Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 5). We examine each of these hypotheses in turn.

@@

To determine category switching, we examine whether an idea is assigned to the same category tree as the previous instance, or whether it was assigned to a different category tree.

We model the probability of any idea category being followed by the same idea category with a Bernoulli random variable. Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (HDI 14.3\% - 17.0\%). To establish a baseline to understand whether this probability is greater than chance, the most popular category tree in our data set is \emph{music player}, representing 5.65\% of all instances. Thus, an upper bound on the greatest probability of an instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI. Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.

@@

To calculate the time to generate a response instance, we examine the times logged for the first and last activation of a text entry widget.

An examination of the timing data suggests a log-normal distribution. We fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.

@@

The resulting models fitted to the data are shown in Figure~\ref{fig:hyp5_comparison}. The mean for the within-category condition was 9.5 in log space (13.4 seconds, HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, HDI 10.2-10.3). The differences in mean are significant and confirm hypothesis 5. Notably, within-category instance generation took an average of 12.6 seconds less than between-categories instance generation, consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between these two idea generation processes.
